---
title: "Assignment 1 - Language Development in ASD - part 2"
author: "[YOUR NAME]"
date: "[DATE]"
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# you could optionally set include = TRUE globally while working on the exercise and then just change it to include = FALSE before you hand in
# knitr::opts_chunk$set(include = FALSE)
```

# Template for the hand-in
### Structure of the code chunks

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format?) and describe the participants (by diagnosis)

We will try to answer three questions:
    
- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries, include = FALSE}
library(ggplot2)
library(lme4)
library(tidyverse)

```

### Define your working directory and load the data
If you created a project for this class and opened this Rmd file from within that project, your working directory is your project directory.

If you opened this Rmd file outside of a project, you will need some code to find the data:
- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

If your're in a project, just put the data in the project folder and you're good to go! (Or make a data subfolder to keep it tidy around here)
```{r Load Data, include = FALSE}
#load data
data <- read_csv("autism_data.csv")

```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

ADOS indicates the severity of the autistic symptoms (the higher the score, the worse the symptoms)
* MLU stands for mean length of utterance
* types stands for unique words (e.g. even if "doggie" is used 100 times it only counts for 1)
* tokens stands for overall amount of words (if "doggie" is used 100 times it counts for 100) 
* MullenRaw indicates non verbal IQ
* ExpressiveLangRaw indicates verbal IQ



```{r descriptive stats, include = FALSE}
#rename columns
names(data)[8]<-"nonverbalIQ"
names(data)[9]<-"verbalIQ"

#create boxplot to check whether the groups are well mathced according to age
boxplot(Age ~ Diagnosis, color = c("white", "ligthgrey"), data ) #the ASD group is in general a little older


#mean age for ASD and TD
data %>%
  group_by(Diagnosis) %>%
  summarise(mean_age = mean(Age, na.rm = TRUE))

#count number of each gender for the two groups
distinct(data, Gender, SUBJ) %>%
  count(Gender == "M") #56 males and 11 females in total
  

distinct(data, Gender, SUBJ, Diagnosis) %>%
  count(Gender == "M", Diagnosis == "ASD") # For ASD, there are 5 females and 26 males. For TD there are 6 females and 30 males.

#more females than males, but the balance between males and females looks similar for the two groups


#plot gender
data %>%
  distinct(Gender, SUBJ, Diagnosis) %>%
  na.omit() %>%
  ggplot(aes(Diagnosis, fill = Gender)) + geom_bar()  #well-matched in relation to gender between the two groups

```

[REPORT THE RESULTS]

## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU (mean length of utterance) changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?

Visits (time)
Diagnosis

- Which variable(s) should be included as random factors?

Subject as random intercept and slope (over visits)

```{r ex2, include = FALSE}
#create MLU_model

MLU_model_1 <- lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ) , data = data, REML = FALSE)

MLU_model_2 <- lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|SUBJ) , data = data, REML = FALSE)

MLU_model_3 <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1+VISIT|SUBJ) , data = data, REML = FALSE) #see if there is an interaction between visit and diagnosis


#nullModel = lmer(CHI_MLU ~ Age + (1|Gender) + (1 + Age|SUBJ), data = AutismData, REML = F)
#nullModel

#both
#ASD.model = lmer(CHI_MLU ~ Age + Diagnosis + (1|Gender) + (1 + Age|SUBJ), data = AutismData, REML = F)
#ASD.model

#interaction
#interactModel = lmer(CHI_MLU ~ Age * Diagnosis + (1|Gender) + (1 + Age|SUBJ), data = AutismData, REML = F)
#interactModel


                   
```

How would you evaluate whether the model is a good model?

```{r ex2 evaluate, include = FALSE}
#use anova to compare the models 

#anova(MLU_model_1, MLU_model_2, MLU_model_3) 

#does not work because there are missing values in one or more of the predictor variables, the default action is to omit all rows that have any values missing

#so have to figure out what we want to with missing data!! 

#we will create new models, which uses the same data (omit rows with no values)


MLU_model_3 <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1+VISIT|SUBJ), data = data, REML = FALSE)


data=na.omit(data[ , all.vars(formula(MLU_model_3))]) #only use the subset data, which were used for making model_3


MLU_model_1 <- lmer(CHI_MLU ~ 1 + VISIT + (1+VISIT|SUBJ), data = data, REML = FALSE)


MLU_model_2 <- lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|SUBJ), data = data, REML = FALSE)



#compare the models with anova
anova(MLU_model_1, MLU_model_2, test="Chisq") #the second model is not significantly better than the first model


anova(MLU_model_1, MLU_model_3, test="Chisq") #model 3 is significantly better than model 1

anova(MLU_model_2, MLU_model_3, test="Chisq") #model 3 is significantly better than model 2

#means that there is an interaction between diagnosis and visit
#indicates that children with ASD has a language impairment


#plot line graph
line_plot <- ggplot(data, aes(VISIT,CHI_MLU, color = Diagnosis)) 

line_plot + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line",aes(group = Diagnosis)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, aes(group = Diagnosis)) + 
  scale_color_manual(values=c("goldenrod1","dodgerblue"))




summary(MLU_model_3)


```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better.


Alternatively you could assume (or test) that language development could be quadratic, that is, its rate of change over time is not constant, but changes either accelerating (the child learns faster and faster, the more language it has to rely on), or decelerating (the child saturates, as there is only that much language it can learn relevant to that specific play situation).

```{r ex2 growth curve, include = FALSE}

line_plot + stat_smooth(method ="lm") #linear

line_plot + stat_smooth(method = "lm", formula = y ~ poly(x,2)) #quadratic

line_plot + stat_smooth(method = "lm", formula = y ~ poly(x,3)) #cubic


```

Quadratic should be something like this: y = a*q + b*q2 + c*q3 


```{r}

#det her virker 
VISIT_quadratic <- lmer(CHI_MLU ~ VISIT + Diagnosis + I(VISIT^2) + (1 + VISIT + I(VISIT^2)|SUBJ), data = data, REML = FALSE)  


#det her virker ikke
VISIT_cubic <- lmer(CHI_MLU ~ VISIT + Diagnosis + I(VISIT^2) +  I(VISIT^3) + (1 + VISIT + I(VISIT^2) + I(VISIT^3)|SUBJ), data = data,REML = FALSE) 

#virker uden cubic on intercept 
VISIT_cubic <- lmer(CHI_MLU ~ VISIT + Diagnosis + I(VISIT^2) +  I(VISIT^3) + (1 + VISIT + I(VISIT^2)|SUBJ), data = data,REML = FALSE) 


#try to rescale

numcols <- grep("^c\\.",names(data))
dfs <- data
dfs[,numcols] <- scale(dfs[,numcols])
new_cubic_model <- update(VISIT_cubic,data=dfs)

#still does not work


```



Exciting right?
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results
- A plot of your best model's predictions

[REPORT THE RESULTS]
Linguistic development of children MLU is affected by ... 

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include = FALSE}

```

[REPORT THE RESULTS]

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Malte (au540041@post.au.dk).


```{r ex4, include = FALSE}

```

[REPORT THE RESULTS]

