---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Riccardo Fusaroli"
date: "August 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
#load libraries
library(tidyverse)
library(ggplot2)
library(lme4)
library(Metrics)
library(caret)



#load data 
train_data <- read.csv("train.csv")

train_data<-train_data[complete.cases(train_data[ , 14]),] #remove na's


test_data <- read.csv("test.csv") #we cleaned the data the same way as we did with the training data in portfolio 1

test_data<-test_data[complete.cases(test_data[ , 14]),] #remove na's


```



## Welcome to the third exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time, then learning how to cross-validate models and finally how to systematically compare models.

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/ra99bdvm6fzay3g/demo_test.csv?dl=1
* Utterance Length data: https://www.dropbox.com/s/uxtqqzl18nwxowq/LU_test.csv?dl=1
* Word data: https://www.dropbox.com/s/1ces4hv8kh0stov/token_test.csv?dl=1

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data () and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the model code again and apply it to your training data (from the first assignment))
- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())
- create the test dataset (apply the code from assignment 1 part 1 to clean up the 3 test datasets)
- test the performance of the models on the test data (Tips: google the functions "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())

formatting tip: If you write code in this document and plan to hand it in, remember to put include=FALSE in the code chunks before handing in.

```{r}

languageModel_base_train = lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=train_data, REML=FALSE) #create model on training data 


predictions_train <- predict(languageModel_base_train, train_data, allow.new.levels = TRUE) #make predictions from model on the training data

rmse_train <-rmse(train_data$CHI_MLU, predictions_train)

rmse_train #use RMSE to calculate the performance of the model on the training data root




#predict on test data
predictions_test <- predict(languageModel_base_train, test_data, allow.new.levels = TRUE)

rmse_test <-rmse(test_data$CHI_MLU, predictions_test) #make predictions from model on the training data

rmse_test #use RMSE to calculate the performance of the model on the training data root


rmse_train
rmse_test

#to get an idea of whether it's better than the error we get by just using the mean as a model
rmse_train/sd(train_data$CHI_MLU)#better than the mean model

rmse_test/sd(test_data$CHI_MLU) #actually worse than the mean model --> we have probably overfitted our model




#RMSE - what is the average error when predicting. A measure of the uncertainty in your prediction

#what is an "okay" error. Divide it by standard deviation in the data. Standard deviation is the deviation around the mean. If the sd is low, when the mean is a good model of the data. 

#we have overfittet on the training set

#a good model used visit*diagnosis + number of unique words (highly correlated with MLU, because they measure the same thing, so if you tell this to the model, it's easy to predict the actual MLU) --> like saying child MLU predicted by child MLU 

#we want to know about language development --> good measures are complexity in speed practices (MLU and number of unique words)


```


[HERE GOES YOUR ANSWER]

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Use cross-validation to compare your model from last week with the basic model (Child MLU as a function of Time and Diagnosis, and don't forget the random effects!)
- (Tips): google the function "createFolds";  loop through each fold, train both models on the other folds and test them on the fold)

Which model is better at predicting new data: the one you selected last week or the one chosen via cross-validation this week?

- Test both of them on the test data.
- Report the results and comment on them.

- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test data



```{r}
#cross validation --> so we do not have to go "throw"  out data in order to test on it 
#make a cake and split it in 5 parts --> 5 folds 
#use cross validation to find the best model without overfitting
#put the kids in a random fold



#create the fancy and the basic model, will just use a difference between linear and a quadratic model

#fancy_model <- lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data= , REML=FALSE)

#basice_model <- lmer(CHI_MLU ~ Diagnosis * VISIT  + (1+VISIT|SUBJ), data= , REML=FALSE)



folds <- createFolds(unique(train_data$SUBJ), k = 5, list = TRUE) #create folds. Unique() puts each unique kid in a fold, so all data from one subj goes into one fold

folds




#now we need to create a loop, important to run through the empty lists before running  
rmse_train_lin = NULL #create empty list

rmse_test_lin = NULL


rmse_train_qua = NULL

rmse_test_qua = NULL

temp_test = NULL

temp_train = NULL



for (f in folds) {
  #divide into test and train
  temp_test = filter(train_data, SUBJ %in% f)
  temp_train = filter(train_data, ! SUBJ %in% f)
  
  #run models on train
  lin_model <- lmer(CHI_MLU ~ Diagnosis * VISIT  + (1+VISIT|SUBJ), data= temp_train, REML=FALSE)
  qua_model <- lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=temp_train, REML=FALSE)
  
  #predict from the two models on both train and test data, train data does not really make sense --> just pedagogic 
  lin_predict_test <- predict(lin_model, temp_test, allow.new.levels = TRUE)
  qua_predict_test <- predict(qua_model, temp_test, allow.new.levels = TRUE)
  lin_predict_train <- predict(lin_model, temp_train, allow.new.levels = TRUE)
  qua_predict_train <- predict(qua_model, temp_train, allow.new.levels = TRUE)
  
  
  #get average error and put into list
  rmse_test_lin <- c(rmse_test_lin, rmse(temp_test$CHI_MLU, lin_predict_test))
  rmse_train_lin <- c(rmse_train_lin, rmse(temp_train$CHI_MLU, lin_predict_train))
  rmse_test_qua <- c( rmse_test_qua, rmse(temp_test$CHI_MLU, qua_predict_test))
  rmse_train_qua <- c(rmse_train_qua, rmse(temp_train$CHI_MLU, qua_predict_train))
  
  
}

#rmse, errors for the linear model on the test data set
 rmse_test_lin
 
 rmse_train_lin
 
 rmse_test_qua
 
 rmse_train_qua
 
 

mean(rmse_test_lin) #get the mean error across the folds for the linear model

mean(rmse_test_qua) #get the mean error across the folds for the quadratic model

mean(rmse_train_lin)

mean(rmse_train_qua)


#run the two models on the test data




```

Which model is better at predicting new data: the one you selected last week or the one chosen via cross-validation this week?

- Test both of them on the test data.
- Report the results and comment on them.

- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test data




```{r}

#test the two models on the test data 

#first train the models
lin_model <- lmer(CHI_MLU ~ Diagnosis * VISIT  + (1+VISIT|SUBJ), data = train_data, REML=FALSE)


qua_model <- lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=train_data, REML=FALSE)


#predict on test data
predictions_test_lin <- predict(lin_model, test_data, allow.new.levels = TRUE)

predictions_test_qua <- predict(qua_model, test_data, allow.new.levels = TRUE)


rmse_test_lin <-rmse(test_data$CHI_MLU, predictions_test_lin) #make predictions from model on the training data

rmse_test_qua <-rmse(test_data$CHI_MLU, predictions_test_qua) #make predictions from model on the training data


rmse_test_lin

rmse_test_qua

#to get an idea of whether it's better than the error we get by just using the mean as a model
rmse_test_lin/sd(train_data$CHI_MLU)#better than the mean model

rmse_test_qua/sd(test_data$CHI_MLU) #actually worse than the mean model --> we have probably overfitted our model

```






### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.
(Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; input the average of the TD group  for each parameter in the model as X1, X2, etc.).

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)


Just to take a look at the equation

First less look at the most simple linear model

The equation is: 

CHI_MLU = 1.046 + VISIT * 0.23 + DIagnosis * 0.29

Diagnosis is either 0 or 1

ASD = 0

TD = 1

```{r}
lin_model_simple <- lmer(CHI_MLU ~ Diagnosis + VISIT  + (1+VISIT|SUBJ), data = train_data, REML=FALSE)

summary(lin_model_simple)
```
Next less look at the next simple linear model --> an interaction model

CHI_MLU = 1.3 + Diagnosis*-0.21 + VISIT*0.1 + Diagnosis*VISIT*0.25

ASD = 0

TD = 1

```{r}
lin_model_interaction <- lmer(CHI_MLU ~ Diagnosis * VISIT  + (1+VISIT|SUBJ), data = train_data, REML=FALSE)

summary(lin_model_interaction)

#CHI_MLU = 1.3 + D* -0.21 + V*0.1 + D*V*0.25

diagnosis=1
visit=1

 1.3 + diagnosis*-0.21 + visit*0.1 + diagnosis*visit*0.25

```

```{r}
qua_model <- lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=train_data, REML=FALSE)

summary(qua_model)

```

CHI_MLU = 0.98 + Diagnosis*-0.36 + VISIT*0.41 + -0.04*VISIT^2 + VISIT*Diagnosis*0.25
```{r}
Diagnosis = 1
Visit = 1


CHI_MLU = 0.98 + Diagnosis*-0.36 + Visit*0.41 + -0.04*Visit^2 + Visit*Diagnosis*0.25

CHI_MLU
average_td = c(1.24, 1.78, 2.24, 2.62, 2.92, 3.14)
tk = average_td - Bernie_data$CHI_MLU
tk

sqrt((tk)^2)

abs(tk)
```



```{r}
#we have anonymized the subject, but we have looked it up and he is participant number 2



#filter Bernie out, don't know if he is participant 1, have to recreate dataset, where we have not removed the names
Bernie_data <- subset(test_data, SUBJ == "2", select = c("SUBJ", "CHI_MLU", "Diagnosis", "VISIT"))

summary(lin_model)



#use the model to predict on data 
predictions_qua_bernie <- predict(qua_model, Bernie_data, allow.new.levels = TRUE)



rmse_qua_bernie <-rmse(Bernie_data$CHI_MLU, predictions_qua_bernie)



rmse_qua_bernie

predictions_qua_bernie


```


[HERE GOES YOUR ANSWER]

### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).

Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.

- Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?

### OPTIONAL: Exercise 5): Using Lasso for model selection

Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!

