---
title: "Assignment 1 - Language Development in ASD - part 2"
author: "Simon M. N."
date: "23/9 2018"
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# you could optionally set include = TRUE globally while working on the exercise and then just change it to include = FALSE before you hand in
# knitr::opts_chunk$set(include = FALSE)
```

# Template for the hand-in
### Structure of the code chunks

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format?) and describe the participants (by diagnosis)

We will try to answer three questions:
    
- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries, include = FALSE}
library(ggplot2)
library(lme4)
library(tidyverse)
library(dplyr)
library(MuMIn)
```

### Define your working directory and load the data
If you created a project for this class and opened this Rmd file from within that project, your working directory is your project directory.

If you opened this Rmd file outside of a project, you will need some code to find the data:
- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

If your're in a project, just put the data in the project folder and you're good to go! (Or make a data subfolder to keep it tidy around here)
```{r Load Data, include = FALSE}
setwd("C:/Users/slmoni/Documents/Uni/Experimental Methods III/R-stuff/Assignment1&2_LanguageASD")
data<-read.csv("ADSFinal_Simon.csv",stringsAsFactors = F)
```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r descriptive stats, include = FALSE}
data<-data[complete.cases(data[ , 14]),]

visit1<- filter(data, VISIT==1)
visit1ASD<- filter(data, VISIT==1 & Diagnosis=='ASD')
visit1TD<- filter(data, VISIT==1 & Diagnosis=='TD')

#Diagnosis
table(visit1$Diagnosis)
#Gender
table(visit1ASD$Gender)
table(visit1TD$Gender)
#Age
t.test(visit1$Age ~ visit1$Diagnosis)
sd(visit1ASD$Age)
sd(visit1TD$Age)
#Ethnicity
table(visit1ASD$Ethnicity)
table(visit1TD$Ethnicity)

#nonverbalIQ
t.test(visit1$nonVerbalIQ ~ visit1$Diagnosis)
#verbalIQ
t.test(visit1$verbalIQ ~ visit1$Diagnosis)
```

[REPORT THE RESULTS]
I first removed the NAs form the list to get a more accurate assessment of the two groups, since some of the children in the dataset only have NAs and therefor have no impact on the analyses. I look only at the children at visit 1 since we are interested in diffrences at the beginning of the experiment and we expect diffrences to occur later in the experiment
First looking at the size of the two groups. Number of children with ASD was 29 and number of children with TD is 31. That is simmilar size groups and we should not run into any issues.
Looking at gender within the ASD group there was 25 females and 4 males and in the TD group there was 25 females and 6 males. Again very similar and it should not cause any issues. One thing to note is that this is not an accurate representation for the real world population, so gernalasation to the real world will be weaker, considering males and females may learn language in diffrenret ways and be effect by ASD in diffrent ways.
A Welch two-sample t-test is run on the average age of the two groups. It was found that there was a significant diffrence between the average age of ASD (M=33, SD=5.59) and the average age of TD (M=20.38, SD=1.51), t(31.82)=11.76 p < 0.001. This may cause some issues since children may learn language at different rates depending on there age. This diffrence is likely due to the difficulty of diagnosing ASD in young children.


## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r ex2, include = FALSE}
boxplot(CHI_MLU ~ Diagnosis*VISIT, col=c("white","lightgray"),data)

line_plot <- ggplot(data, aes(VISIT,CHI_MLU, color = Diagnosis)) 
line_plot + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line",aes(group = Diagnosis)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, aes(group = Diagnosis)) + 
  scale_color_manual(values=c("goldenrod1","dodgerblue"))

languageModel = lmer(CHI_MLU ~ Diagnosis * VISIT + (1+VISIT|SUBJ), data=data,REML=FALSE)
summary(languageModel)
```

How would you evaluate whether the model is a good model?

```{r ex2 evaluate, include = FALSE}
r.squaredGLMM(languageModel)
languageModel = lmer(CHI_MLU ~ Diagnosis * VISIT + (1+VISIT|SUBJ), data=data, REML=FALSE)
languageNull = lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ), data=data, REML=FALSE)
anova(languageNull,languageModel)
```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better

```{r ex2 growth curve, include = FALSE}
languageModel2 = lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModel2)
r.squaredGLMM(languageModel2)
languageNull2 = lmer(CHI_MLU ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageNull2,languageModel2)
```

Exciting right?
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results
- A plot of your best model's predictions

[REPORT THE RESULTS]
Linguistic development of children MLU is affected by ... 

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include = FALSE}
boxplot(MOT_MLU ~ Diagnosis*VISIT, col=c("white","lightgray"),data)

line_plot <- ggplot(data, aes(VISIT,MOT_MLU, color = Diagnosis)) 

line_plot + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line",aes(group = Diagnosis)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, aes(group = Diagnosis)) + 
  scale_color_manual(values=c("goldenrod1","dodgerblue"))

languageModelMot = lmer(MOT_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModelMot)
r.squaredGLMM(languageModelMot)
languageNullMot = lmer(MOT_MLU ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageNullMot,languageModelMot)
```

[REPORT THE RESULTS]

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Malte (au540041@post.au.dk).


```{r ex4, include = FALSE}
boxplot(types_MOT ~ Diagnosis*VISIT, col=c("white","lightgray"),data)
languageModelTest = lmer(types_MOT ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModelTest)
languageModelTestNull = lmer(types_MOT ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageModelTestNull,languageModelTest)
```

[REPORT THE RESULTS]

