---
title: "Assignment 1 - Language Development in ASD - part 2"
author: "Studiegroup"
date: "25/9 2018"
output:
  word_document: default
  html_document: default
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# you could optionally set include = TRUE globally while working on the exercise and then just change it to include = FALSE before you hand in
# knitr::opts_chunk$set(include = FALSE)
```

# Template for the hand-in
### Structure of the code chunks

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format?) and describe the participants (by diagnosis)

We will try to answer three questions:
    
- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries, include = FALSE}
#Loading packages
library(ggplot2)
library(effects)
library(lme4)
library(tidyverse)
library(dplyr)
library(MuMIn)
library(Hmisc)
```

### Define your working directory and load the data
If you created a project for this class and opened this Rmd file from within that project, your working directory is your project directory.

If you opened this Rmd file outside of a project, you will need some code to find the data:
- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

If your're in a project, just put the data in the project folder and you're good to go! (Or make a data subfolder to keep it tidy around here)
```{r Load Data, include = FALSE}
# Setting WD and loading the data

data<-read.csv("ADSFinal_Simon.csv",stringsAsFactors = F)

# Changing diagnosis to a factor to avoid issues later
data$Diagnosis <-as.factor(data$Diagnosis)
```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r descriptive stats, include = FALSE}
# Removing NA's
data<-data[complete.cases(data[ , 14]),]

# Making three dataframs with only the first visit and for only the ASD and TD group
visit1<- filter(data, VISIT==1)
visit1ASD<- filter(data, VISIT==1 & Diagnosis=='ASD')
visit1TD<- filter(data, VISIT==1 & Diagnosis=='TD')

#Diagnosis comparison
table(visit1$Diagnosis)
#Gender comparison
table(visit1ASD$Gender)
table(visit1TD$Gender)
#Age comparison
t.test(visit1$Age ~ visit1$Diagnosis)
sd(visit1ASD$Age)
sd(visit1TD$Age)
#Ethnicity comparison
table(visit1ASD$Ethnicity)
table(visit1TD$Ethnicity)
#nonverbalIQ comparison
t.test(visit1$nonVerbalIQ ~ visit1$Diagnosis)
sd(visit1ASD$nonVerbalIQ)
sd(visit1TD$nonVerbalIQ)
#verbalIQ comparison
t.test(visit1$verbalIQ ~ visit1$Diagnosis)
sd(visit1ASD$verbalIQ)
sd(visit1TD$verbalIQ)
```

The data is divided into two groups: (1) children with the ASD diagnosis (2) children with the TD diagnosis (Typically Developing is in medical terms not really a diagnosis, but it will for practical reasons be described as such in the following texts). 

Upon first inspecting the data, two initial issues raise concern. First, the unequal distribution of gender (Males are heavily favored) may not accurately reflect the population which this study hopes to generalize towards. This may be a problem if males and females are affected by ASD in different ways. However, the distribution of genders does seem somewhat similar across Diagnosis, indicating that the groups are comparable.

Secondly, the age difference between the ASD and TD groups is significant (t=11.761, p < 0.05). This could be a result of the generally slower development of linguistic capabilities in  ASD diagnosed children. Since the participants are selected based on their linguistic performance at visit 1, we can assume that ASD diagnosed children are generally older before they reach the same linguistic capabilities as the TD diagnosed children. Supporting the notion that ASD diagnosed children are slower in their development, is the result that there is no notable difference in the verbalIQ (p > 0.05). These finds support the conclusion that the ASD group, though significantly older than the TD group, have not significantly further developed linguistic capabilities. 

Other noteworthy observations include: 
  - Similar sized groups across diagnosis (29 ASD, 31 TD)
  - No singificant difference in nonverbal-IQ (p > 0.05)



## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r ex2, echo= FALSE}
###Working with childs data
#Making a boxplot of the data to assess the data
boxplot(CHI_MLU ~ Diagnosis*VISIT, col=c("white","lightgray"),data)

#Makeing a lineplot of the data to assess the data
line_plot <- ggplot(data, aes(VISIT,CHI_MLU, color = Diagnosis)) 
line_plot + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line",aes(group = Diagnosis)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, aes(group = Diagnosis)) + 
  scale_color_manual(values=c("goldenrod1","dodgerblue"))

#Making a simple linear mixed effects model. Model 1
languageModel = lmer(CHI_MLU ~ Diagnosis * VISIT + (1+VISIT|SUBJ), data=data,REML=FALSE)
summary(languageModel)
```

At first glance the visualizations of the model indicates a significant difference in the development of MLU over time between TD and ASD diagnosed children, with the TD children increasing their MLU more than the ASD children. 

How would you evaluate whether the model is a good model?

```{r ex2 evaluate, include = FALSE}
## Assessing model 1

# First with r squared
r.squaredGLMM(languageModel)

# Then with an anova
languageNull = lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ), data=data, REML=FALSE)
anova(languageNull,languageModel)

#Assumptions testing
plot(residuals(languageModel))
qqnorm(residuals(languageModel))
plot(fitted(languageModel),residuals(languageModel))^2
```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better

```{r ex2 growth curve, include= FALSE}
#Making a quadratic mixed effects model. Model 2
languageModel2 = lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModel2)
# Assessing model 2
r.squaredGLMM(languageModel2)
languageNull2 = lmer(CHI_MLU ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageNull2,languageModel2)

#Making a cubic mixed effects model. Model 3
model_cubic <- lmer(CHI_MLU ~ VISIT * Diagnosis + I(VISIT^2) + I(VISIT^3)  + (1+VISIT + I(VISIT^2) + I(VISIT^3)|SUBJ), data = data, REML=FALSE)
# Assessing model 3
summary(model_cubic)
r.squaredGLMM(model_cubic)

#Making a "what ever this is called" mixed effects model. Model 4
model_quadrouble <- lmer(CHI_MLU ~ VISIT * Diagnosis + I(VISIT^2) + I(VISIT^3) +I(VISIT^4)+ (1+VISIT + I(VISIT^2) + I(VISIT^3) + I(VISIT^4)|SUBJ), data = data, REML=FALSE)
# Assessing model 4
summary(model_quadrouble)
r.squaredGLMM(model_quadrouble)

# Plot of the child MLU model (shamelessly stolen with comments and all :P)
ee <- effect(c("Diagnosis","VISIT"),languageModel2) 
theme_set(theme_bw())
ggplot(as.data.frame(ee),
       aes(VISIT,fit,colour=Diagnosis,fill=Diagnosis))+
    geom_line()+
     ## colour=NA suppresses edges of the ribbon
    geom_ribbon(colour=NA,alpha=0.1,
                            aes(ymin=lower,ymax=upper))+
  labs(title = "Diagnosis as predictor of child MLU development", x = "Visit", y = "MLU")

#Assumptions testing
plot(residuals(languageModel2))
qqnorm(residuals(languageModel2))
plot(fitted(languageModel2),residuals(languageModel2))^2

plot(residuals(model_cubic))
qqnorm(residuals(model_cubic))
plot(fitted(model_cubic),residuals(model_cubic))^2

plot(residuals(model_quadrouble))
qqnorm(residuals(model_quadrouble))
plot(fitted(model_quadrouble),residuals(model_quadrouble))^2
```

```{r, echo= FALSE}
# Plot of the child MLU model (shamelessly stolen with comments and all :P)
ee <- effect(c("Diagnosis","VISIT"),languageModel2) 
theme_set(theme_bw())
ggplot(as.data.frame(ee),
       aes(VISIT,fit,colour=Diagnosis,fill=Diagnosis))+
    geom_line()+
     ## colour=NA suppresses edges of the ribbon
    geom_ribbon(colour=NA,alpha=0.1,
                            aes(ymin=lower,ymax=upper))+
  labs(title = "Diagnosis as predictor of child MLU development", x = "Visit", y = "MLU")

qqnorm(residuals(languageModel2))

```

Exciting right?
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results
- A plot of your best model's predictions

Linguistic development of children MLU is affected by diagnosis. Results indicate that children with the ASD diagnosis increase in MLU more slowly than the control group. 


Comparing between a quadratic, cubic, and a quadrouble model, the quadratic model best explained the fixed effects of the quadratic model was the one which best explained the variance. This suggests that the learning curve describing increase in MLU over time (visits to be accurate) is best fitted to a parabola. 


A quadratic mixed effect analysis was performed to determine the impact of ASD on the linguistic development (measured in mean length of utterances (MLU)) in children over a period of 20 months (6 visits with 4 months apart). As we are interested in examining the different developments of ASD and TD children  over time, the fixed effects set as the interaction effect between Diagnosis (ASD|TD) and number of Visit. As we expect each child to have an individual learning curve and starting point, the random effects were intercepts and slopes set for subjects. The results (bata=0.25, SE=0.04, t(12)=6.73, p < 0.001) indicate that difference in diagnosis significantly predicts the development of child MLU to such an extent that MLU increases by 0.25 more pr. visit for children with the TD diagnosis. 

We tested for assumptions of normality and homoscedasticity and they were not violated

In relation to our hypothesis, the null-hypothesis must be: "i) Child MLU does not change over time, ii) according to diagnosis". Based on our results we can reject the null-hypothesis.

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include FALSE}
###Working with mothers data
#Making a boxplot of the data to assess the data
boxplot(MOT_MLU ~ Diagnosis*VISIT, col=c("white","lightgray"),data)

#Makeing a lineplot of the data to assess the data
line_plot <- ggplot(data, aes(VISIT,MOT_MLU, color = Diagnosis)) 
line_plot + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line",aes(group = Diagnosis)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, aes(group = Diagnosis)) + 
  scale_color_manual(values=c("goldenrod1","dodgerblue"))

#Making a quadratic mixed effects model. Model 5
languageModelMot = lmer(MOT_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModelMot)
# Assessing model 4
r.squaredGLMM(languageModelMot)
languageNullMot = lmer(MOT_MLU ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageNullMot,languageModelMot)

# Plot of the parent MLU model (shamelessly stolen with comments and all :P)
ee <- effect(c("Diagnosis","VISIT"),languageModelMot) 
theme_set(theme_bw())
ggplot(as.data.frame(ee),
       aes(VISIT,fit,colour=Diagnosis,fill=Diagnosis))+
    geom_line()+
     ## colour=NA suppresses edges of the ribbon
    geom_ribbon(colour=NA,alpha=0.1,
                            aes(ymin=lower,ymax=upper))+
  labs(title = "Diagnosis as predictor of mother MLU increase", x = "Visit", y = "MLU") 

#Assumptions testing
plot(residuals(languageModelMot))
qqnorm(residuals(languageModelMot))
plot(fitted(languageModelMot),residuals(languageModelMot))^2
```

```{r, echo= FALSE}

# Plot of the parent MLU model (shamelessly stolen with comments and all :P)
ee <- effect(c("Diagnosis","VISIT"),languageModelMot) 
theme_set(theme_bw())
ggplot(as.data.frame(ee),
       aes(VISIT,fit,colour=Diagnosis,fill=Diagnosis))+
    geom_line()+
     ## colour=NA suppresses edges of the ribbon
    geom_ribbon(colour=NA,alpha=0.1,
                            aes(ymin=lower,ymax=upper))+
  labs(title = "Diagnosis as predictor of mother MLU increase", x = "Visit", y = "MLU")

qqnorm(residuals(languageModelMot))

```

Parental MLU over time is affected by the diagnosis of the belonging child. 

A quadratic mixed effect analysis was performed to determine the impact of ASD on the linguistic development (measured in mean length of utterances (MLU)) in mothers communication to their children over a period of 20 months (6 visits with 4 months apart). As we are interested in examining the different developments in communication between mothers of ASD and TD children over time, the fixed effects set as the interaction effect between Diagnosis (ASD|TD) and number of Visit. As we expect each mother to have an individual learning curve and starting point, the random effects were intercepts and slopes set for subjects. The results (bata=0.03, SE=0.035, t(12)=0.857, p < 0.001) indicate that difference in diagnosis significantly predicts the development of Mother MLU to such an extent that MLU increases by 0.03 more pr. visit for mothers with children with the TD diagnosis. This effect seems noticeably smaller than the increase of child MLU. The small (BUT SIGNIFICANT) effect begs the quesition of whether age has a larger effect on the mothers MLU than diagnosis. 

We tested for assumptions of normality and homoscedasticity and they were not violated

In relation to our hypothesis, the null-hypothesis must be: "i) Mother MLU does not change over time, ii) according to diagnosis". Based on our results we can reject the null-hypothesis.

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Malte (au540041@post.au.dk).

```{r ex4, include = FALSE}
#You know the deal by now :|
boxplot(types_CHI ~ Diagnosis*VISIT, col=c("white","lightgray"),data)
languageModelTest = lmer(types_CHI ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModelTest)
r.squaredGLMM(languageModelTest)
languageModelTestNull = lmer(types_MOT ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageModelTestNull,languageModelTest)
```

We explored the data to find better models, but we were unable to. 
Nothing more too see here

















:3