---
title: "Assignment 1 - Language Development in ASD - part 2"
author: "Simon M. N."
date: "25/9 2018"
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# you could optionally set include = TRUE globally while working on the exercise and then just change it to include = FALSE before you hand in
# knitr::opts_chunk$set(include = FALSE)
```

# Template for the hand-in
### Structure of the code chunks

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format?) and describe the participants (by diagnosis)

We will try to answer three questions:
    
- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries, include = FALSE}
#Loading packages
library(ggplot2)
library(lme4)
library(tidyverse)
library(dplyr)
library(MuMIn)
library(effects)
```

### Define your working directory and load the data
If you created a project for this class and opened this Rmd file from within that project, your working directory is your project directory.

If you opened this Rmd file outside of a project, you will need some code to find the data:
- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

If your're in a project, just put the data in the project folder and you're good to go! (Or make a data subfolder to keep it tidy around here)
```{r Load Data, include = FALSE}
# Setting WD and loading the data
setwd("C:/Users/slmoni/Documents/Uni/Experimental Methods III/R-stuff/Assignment1&2_LanguageASD")
data<-read.csv("ADSFinal_Simon.csv",stringsAsFactors = F)

# Changing diagnosis to a factor to avoid issues later
data$Diagnosis <-as.factor(data$Diagnosis)
```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r descriptive stats, include = FALSE}
# Removing NA's
data<-data[complete.cases(data[ , 14]),]

# Making three dataframs with only the first visit and for only the ASD and TD group
visit1<- filter(data, VISIT==1)
visit1ASD<- filter(data, VISIT==1 & Diagnosis=='ASD')
visit1TD<- filter(data, VISIT==1 & Diagnosis=='TD')

#Diagnosis comparison
table(visit1$Diagnosis)
#Gender comparison
table(visit1ASD$Gender)
table(visit1TD$Gender)
#Age comparison
t.test(visit1$Age ~ visit1$Diagnosis)
sd(visit1ASD$Age)
sd(visit1TD$Age)
#Ethnicity comparison
table(visit1ASD$Ethnicity)
table(visit1TD$Ethnicity)
#nonverbalIQ comparison
t.test(visit1$nonVerbalIQ ~ visit1$Diagnosis)
sd(visit1ASD$nonVerbalIQ)
sd(visit1TD$nonVerbalIQ)
#verbalIQ comparison
t.test(visit1$verbalIQ ~ visit1$Diagnosis)
sd(visit1ASD$verbalIQ)
sd(visit1TD$verbalIQ)
```

[REPORT THE RESULTS]

## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r ex2, include = FALSE}
###Working with childs data
#Making a boxplot of the data to assess the data
boxplot(CHI_MLU ~ Diagnosis*VISIT, col=c("white","lightgray"),data)

#Makeing a lineplot of the data to assess the data
line_plot <- ggplot(data, aes(VISIT,CHI_MLU, color = Diagnosis)) 
line_plot + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line",aes(group = Diagnosis)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, aes(group = Diagnosis)) + 
  scale_color_manual(values=c("goldenrod1","dodgerblue"))

#Making a simple linear mixed effects model. Model 1
languageModel = lmer(CHI_MLU ~ Diagnosis * VISIT + (1+VISIT|SUBJ), data=data,REML=FALSE)
summary(languageModel)
```

How would you evaluate whether the model is a good model?

```{r ex2 evaluate, include = FALSE}
## Assessing model 1

# First with r squared
r.squaredGLMM(languageModel)

# Then with an anova
languageNull = lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ), data=data, REML=FALSE)
anova(languageNull,languageModel)

#Assumptions testing
plot(residuals(languageModel))
qqnorm(residuals(languageModel))
plot(fitted(languageModel),residuals(languageModel))^2
```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better

```{r ex2 growth curve, include = FALSE}
#Making a quadratic mixed effects model. Model 2
languageModel2 = lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModel2)
# Assessing model 2
r.squaredGLMM(languageModel2)
languageNull2 = lmer(CHI_MLU ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageNull2,languageModel2)

#Making a cubic mixed effects model. Model 3
model_cubic <- lmer(CHI_MLU ~ VISIT * Diagnosis + I(VISIT^2) + I(VISIT^3)  + (1+VISIT + I(VISIT^2) + I(VISIT^3)|SUBJ), data = data, REML=FALSE)
# Assessing model 3
summary(model_cubic)
r.squaredGLMM(model_cubic)

#Making a "what ever this is called" mixed effects model. Model 4
model_quadrouble <- lmer(CHI_MLU ~ VISIT * Diagnosis + I(VISIT^2) + I(VISIT^3) +I(VISIT^4)+ (1+VISIT + I(VISIT^2) + I(VISIT^3) + I(VISIT^4)|SUBJ), data = data, REML=FALSE)
# Assessing model 4
summary(model_quadrouble)
r.squaredGLMM(model_quadrouble)

# Plot of the child MLU model (shamelessly stolen with comments and all :P)
ee <- effect(c("Diagnosis","VISIT"),languageModel2) 
theme_set(theme_bw())
ggplot(as.data.frame(ee),
       aes(VISIT,fit,colour=Diagnosis,fill=Diagnosis))+
    geom_line()+
     ## colour=NA suppresses edges of the ribbon
    geom_ribbon(colour=NA,alpha=0.1,
                            aes(ymin=lower,ymax=upper))+
  labs(title = "Diagnosis as predictor of child MLU development", x = "Visit", y = "MLU")

#Assumptions testing
plot(residuals(languageModel2))
qqnorm(residuals(languageModel2))
plot(fitted(languageModel2),residuals(languageModel2))^2

plot(residuals(model_cubic))
qqnorm(residuals(model_cubic))
plot(fitted(model_cubic),residuals(model_cubic))^2

plot(residuals(model_quadrouble))
qqnorm(residuals(model_quadrouble))
plot(fitted(model_quadrouble),residuals(model_quadrouble))^2
```

Exciting right?
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results
- A plot of your best model's predictions
Linguistic development of children MLU is affected by ... 

[REPORT THE RESULTS]

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include = FALSE}
###Working with mothers data
#Making a boxplot of the data to assess the data
boxplot(MOT_MLU ~ Diagnosis*VISIT, col=c("white","lightgray"),data)

#Makeing a lineplot of the data to assess the data
line_plot <- ggplot(data, aes(VISIT,MOT_MLU, color = Diagnosis)) 
line_plot + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line",aes(group = Diagnosis)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, aes(group = Diagnosis)) + 
  scale_color_manual(values=c("goldenrod1","dodgerblue"))

#Making a quadratic mixed effects model. Model 5
languageModelMot = lmer(MOT_MLU ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModelMot)
# Assessing model 4
r.squaredGLMM(languageModelMot)
languageNullMot = lmer(MOT_MLU ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageNullMot,languageModelMot)

# Plot of the parent MLU model (shamelessly stolen with comments and all :P)
ee <- effect(c("Diagnosis","VISIT"),languageModelMot) 
theme_set(theme_bw())
ggplot(as.data.frame(ee),
       aes(VISIT,fit,colour=Diagnosis,fill=Diagnosis))+
    geom_line()+
     ## colour=NA suppresses edges of the ribbon
    geom_ribbon(colour=NA,alpha=0.1,
                            aes(ymin=lower,ymax=upper))+
  labs(title = "Diagnosis as predictor of mother MLU increase", x = "Visit", y = "MLU") 

#Assumptions testing
plot(residuals(languageModelMot))
qqnorm(residuals(languageModelMot))
plot(fitted(languageModelMot),residuals(languageModelMot))^2
```

[REPORT THE RESULTS]

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Malte (au540041@post.au.dk).

```{r ex4, include = FALSE}
#You know the deal by now :|
boxplot(types_CHI ~ Diagnosis*VISIT, col=c("white","lightgray"),data)
languageModelTest = lmer(types_CHI ~ Diagnosis * VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
summary(languageModelTest)
r.squaredGLMM(languageModelTest)
languageModelTestNull = lmer(types_MOT ~ VISIT + I(VISIT^2) + (1+VISIT+ I(VISIT^2)|SUBJ), data=data, REML=FALSE)
anova(languageModelTestNull,languageModelTest)
```

[REPORT THE RESULTS]
Nothing more too see here

















:3