---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Laurits Dixen"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(caret)
library(Metrics)
library(lmerTest)
library(boot)
library(pROC)

#loading the data
data = read.csv("schizoAcousticData.csv")
# Making the subjects paired again
data$Subject = ifelse(data$Diagnosis==0, data$Subject-1000, data$Subject) 

#making list for feature names for later use
features = colnames(data)[6:16]

```

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

### Question 1: Can you diagnose schizophrenia from pitch range only? If so, how well?

Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

```{r, include=FALSE}

#turning pitch range into z-score (needed for the model to converge) (can also be done like this: data$z.qr95 = scale(data$qr95, center = TRUE, scale = TRUE))
data$z.qr95 = (data$qr95 - mean(na.omit(data$qr95))) / sd(na.omit(data$qr95))

#create a model with diagnosis predicted by pitch range (0.5 to 0.95 quantiles) and random effects
model = glmer(Diagnosis ~ z.qr95 + (1|Triangle) + (1|Subject), family = "binomial",data = data)
summary(model)

#creating extra columns with all features z-scored 
for (f in features){ #loop that runs through the list of features and paste it into the calculation
  eval(parse(text = paste("data$z.",f," = (data$",f," - mean(na.omit(data$",f,"))) / sd(na.omit(data$",f,"))", sep="")))
}

#making list with z-scored feature names for later use
zfeatures = c(colnames(data)[32:40],"z.qr95")

#add a column with predictions in the unit "probabilities" (with inv.logit)
data$predictions = inv.logit(predict(model, newdata = data, allow.new.levels=TRUE))

#set threshold for diagnosing at 0.5
data$predictions[data$predictions > 0.5] = "Schizo"
data$predictions[data$predictions <= 0.5] = "Control"

#align labels of diagnosis
data$Diagnosis[data$Diagnosis == 1] = "Schizo"
data$Diagnosis[data$Diagnosis == 0] = "Control"

#create confusion matrix to get performance measure
confusionMatrix(data = as.factor(data$predictions), reference = as.factor(data$Diagnosis), positive = "Schizo")

#load predictions again to get numeric data
data$predictions = inv.logit(predict(model, newdata = data, allow.new.levels=TRUE))

#ROC curve
rocCurve <- roc(response = data$Diagnosis, predictor= data$predictions)
#get area under the curve
auc(rocCurve)
#95% confidence intervals
ci(rocCurve)
#plot the curve
plot(rocCurve, legacy.axes=TRUE)

```


Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

```{r, include=FALSE}
#make diagnosis numeric again
data$Diagnosis[data$Diagnosis == "Schizo"] = 1
data$Diagnosis[data$Diagnosis == "Control"] = 0
data$Diagnosis = as.numeric(data$Diagnosis)

#cross validation with 5 folds over the pitch range model
folds = createFolds(y = unique(data$Subject), k=5, list = TRUE, returnTrain = FALSE)
folds = lapply(folds, function(x) unique(data$Subject)[x])
testSet = NULL
for (f in folds) {
  test = filter(data, Subject %in% f)
  train = filter(data, !Subject %in% f)
  
  tempModel = glmer(Diagnosis ~ z.qr95 + (1|Subject) + (1|Triangle), family = "binomial",data = train)
  
  test$predictions = inv.logit(predict(tempModel, newdata = test, allow.new.levels=TRUE))

  # Creating the full dataframe of predictions (can also be done like this: "predictions[data$Subject %in% f] = test$predictions", need to create emtpy vector before loop, like this "predictions = rep(NA, nrow(data))")
  if (!is.null(testSet)){
    testSet = rbind(testSet,test)
  } else {
    testSet = test
  }  
}

#set threshold for diagnosing at 0.5 and change label
testSet$predictions[testSet$predictions > 0.5] = "Schizo"
testSet$predictions[testSet$predictions <= 0.5] = "Control"

#align labels of diagnosis
testSet$Diagnosis[testSet$Diagnosis == 1] = "Schizo"
testSet$Diagnosis[testSet$Diagnosis == 0] = "Control"

#create confusion matrix to get performance measures
confusionMatrix(data = as.factor(testSet$predictions), reference = as.factor(testSet$Diagnosis), positive = "Schizo")

```


N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

### Question 2 - Which single acoustic predictor is the best predictor of diagnosis?

```{r, include=FALSE}
#make diagnosis numeric again
data$Diagnosis[data$Diagnosis == "Schizo"] = 1
data$Diagnosis[data$Diagnosis == "Control"] = 0
data$Diagnosis = as.numeric(data$Diagnosis)

#cross validate the models (one for each acoustic predictor)
folds = createFolds(y = unique(data$Subject), k=5, list = TRUE, returnTrain = FALSE)
folds = lapply(folds, function(x) unique(data$Subject)[x])
testSet = NULL 
for (f in folds) {
  test = filter(data, Subject %in% f)
  train = filter(data, !Subject %in% f)
  
  for (zf in zfeatures) { #loop that runs through the list of z.features  
    eval(parse(text=paste('tempModel = glmer(Diagnosis ~ ',zf,' + (1|Subject) + (1|Triangle), family = "binomial",data = train)', sep='' )))
    eval(parse(text=paste('test$pred',zf,' = inv.logit(predict(tempModel, newdata = test, allow.new.levels=TRUE))',sep='' )))
    eval(parse(text=paste('test$pred',zf,'[test$pred',zf,' > 0.5] = "Schizo"',  sep=''))) #set threshold
    eval(parse(text=paste('test$pred',zf,'[test$pred',zf,' <= 0.5] = "Control"',sep='')))
  }
  
  # Creating the full dataframe of predictions
  if (!is.null(testSet)){
    testSet = rbind(testSet,test)
  } else {
    testSet = test
  }  
}

#change labels for diagnosis
testSet$Diagnosis[testSet$Diagnosis == 1] = "Schizo"
testSet$Diagnosis[testSet$Diagnosis == 0] = "Control"

#make confusion matrix to get performance measures
for (zf in zfeatures){ #loop that runs through all the z.features 
  eval(parse(text=paste('matrix = confusionMatrix(data = as.factor(testSet$pred',zf,'), reference = as.factor(testSet$Diagnosis), positive = "Schizo")',sep = '')))
  print(paste(zf,"has an accuracy of:",round(matrix$overall[1],3), "a sensitivity of:" ,round(matrix$byClass[1],3), "and a specificity of:", round(matrix$byClass[2], 3))) #print the accuracy score for all the models
}



```


### Question 3 - Which combination of acoustic predictors is best for diagnosing schizophrenia?

Now it's time to go wild! Use all (voice-related) variables and interactions you can think of. Compare models and select the best performing model you can find.

Remember:
- Out-of-sample error crucial to build the best model!
- After choosing the model, send Malte and Riccardo the code of your model


```{r}


windowsPath = "/Users/lauritsdixen/R-stuff/Assignment 2/cogsci-methods-3-assignments-master/Assignment3_VoiceSchizo/Pitch/" #set path to the palce you saved the data



read_pitch <- function(filename) {
    # read data
  data <- read.delim(file = paste(windowsPath, filename, sep = ""), sep = "\t", header = TRUE) 
  soundname = filename
    # parse filename and extract metadata
  meta = str_extract_all(filename, "\\d+")
  study = as.numeric(meta[[1]][1])
  Diagnosis = as.numeric(meta[[1]][2])
  Subject = as.numeric(meta [[1]][3])
  triangles = as.numeric(meta[[1]][4])
  
    # extract descriptors
  mean_f0 = round(mean(data$f0), digits = 2)
  sd = sd(data$f0)
  quantiles = quantile(data$f0, probs = c(0.05, 0.25, 0.50, 0.75, 0.95))
  q05 = quantiles[[1]]
  q25 = quantiles[[2]]
  median = quantiles[[3]]
  q75 = quantiles[[4]]
  q95 = quantiles[[5]]
  iqr2575 = q75-q25
  coe.var = coefficient.variation(sd(data$f0), mean(data$f0))
  MAD = mad(data$f0) #average distance from something...
  
    # combine all this data
  output_list = data.frame(Subject, Diagnosis, study, triangles, mean_f0, sd, q05, q25, median, q75, q95, iqr2575, coe.var, MAD,soundname)
  
  return(output_list)
}

#create dataframe with collected features and metadata
pitch_data = list.files(path = windowsPath) %>%
    purrr::map_df(read_pitch)

#read articulation data
extra = read.delim("Articulation.txt", sep=",")
extra$soundname = as.character(extra$soundname)
pitch_data$soundname = str_replace(pitch_data$soundname,"_f0.txt","")

#merge articulation data with pitch data
fullData = full_join(pitch_data,extra)
fullData = fullData[complete.cases(fullData[ , 1]),]

#make list with column names for all features, both articulation and pitch
newFeatures = c(colnames(fullData)[5:14],colnames(fullData)[16:22])

#rescale features, by z-scoring and min/max 
for (f in newFeatures){
  eval(parse(text = paste("fullData$z.",f," = (fullData$",f," - mean(na.omit(fullData$",f,"))) / sd(na.omit(fullData$",f,"))", sep="")))
  eval(parse(text = paste("fullData$m.",f," = (fullData$",f,"-min(na.omit(fullData$",f,"))) /(max(na.omit(fullData$",f,"))-min(na.omit(fullData$",f,")))",sep="")))
}
```



```{r}
# Defining possible variables, choose which normalization (z-score or min/max)
#variables <- c('m.iqr2575','m.sd','m.MAD','m.coe.var','m.npause','m.dur..s.')
variables <- c('z.iqr2575','z.sd','z.MAD','z.coe.var','z.npause','z.dur..s.')

#define signs to be able to later specify whether it's and interaction or not
signs = list(c("+","*"))

#create empty dataframe forholding performance measures of the model
models = data.frame(matrix(ncol=4),nrow=0)
colnames(models) = c("variables", "accuracy", "sensi", "speci","AUC")

#create folds for cross validation
folds = createFolds(y = unique(fullData$Subject), k=5, list = TRUE, returnTrain = FALSE)
folds = lapply(folds, function(x) unique(fullData$Subject)[x])
# Creating a powerset (set of all possible subsets)
PowerSet <- powerSet(variables)
for (set in PowerSet[2:length(PowerSet)]) {
  # Reformating the set
  if (length(set)==1) {
    collapsed = paste(set)
  } else {
    currentSet = c()
    for (e in set) {
      currentSet <- c(e,currentSet) }
    #dataframe with all possible combinations of signs
    signsCombs = expand.grid(rep(signs, length(currentSet)-1))
    
    for (row in seq(1,nrow(signsCombs))){
      combined = c(currentSet[1])
      
      for (col in seq(2,length(currentSet))) {
        sign = as.character(signsCombs[row,col-1])
        combined = c(combined,sign,currentSet[col])
      }
      collapsed = paste(combined, collapse ="") #collapse into one string 
      print(collapsed)
      # Creating the command in string format
      testParsing <- paste('glmer(Diagnosis ~ ',collapsed,' + (1|Subject) + (1|triangles), data = train, family = "binomial")',sep = '')
      predictors = rep(NA, nrow(fullData)) #create empty list to save predicitons in 
      # CV folds
      for (f in folds) {
        test = filter(fullData, Subject %in% f)
        train = filter(fullData, !Subject %in% f)
        #keep track of error
        ERR <<- FALSE
        #if cathcing an error for one of the model, stop cross validation for the model, move on to next model
        tryCatch ({model = eval(parse(text=testParsing))},
        error=function(err){ERR <<- TRUE; print ("Not Converged")}) 
        if (ERR) {
          break
        } 
        # Creating predictions
        predictors[fullData$Subject %in% f] = as.numeric(inv.logit(predict(model, newdata = test, allow.new.levels=TRUE)))
      }
      if (!ERR) { #only make predictions if error is false (model could converge)
        predictors = as.numeric(predictors)
        rocCurve <- roc(response = fullData$Diagnosis, predictor = predictors)
        fullData$Predictors = predictors
        #take the average of prediction for each subject
        dat = aggregate(fullData[,"Predictors"],list(fullData$Subject,fullData$Diagnosis),mean,na.rm=TRUE)

        colnames(dat) = c("Subject","Diagnosis","Predictors")
        dat$Predictors = ifelse(dat$Predictors > 0.5, 1, 0) #for each subject, define threshold
        matrix = confusionMatrix(data = as.factor(dat$Predictors), reference = as.factor(dat$Diagnosis),positive = "1")
        models = rbind(models, c(collapsed,matrix$overall[1],matrix$byClass[1],matrix$byClass[2],auc(rocCurve))) #save performance measure
      }
    }
  }
}
```





### Question 4: Properly report the results

METHODS SECTION: how did you analyse the data? That is, how did you extract the data, designed the models and compared their performance?

#QUESTION 1: How did we analyse the data? 

From the last portfolio we had a file with extraced acoustic features. From this file we used the acoustic feature "qr95", which is the pitch range from quartile 0.05 to quartile 0.95. This range was used in order to avoid outliers to drastically affect the acoustic feature.

We z-scored all of the acoustic feature, since an assumption when making the models is that the continious variables are on a somewhat similar scale. So if one variable is on a scale from 0 to 1000 and another variable on a scale from 1 to 3 the model fails to converge.


Model design: 

The model used to investigate wheter pitch range is a significant predictor of diagnosis was as follows: 

model = glmer(Diagnosis ~ range + (1|Triangle) + (1|Subject), family = "binomial")

We made a logistic regression, since the outcome we want to predict is binomial, i.e. Schizophrenia or Control. 
We included Subject and Triangle as random intercept. This is to allow subjects to have different pitch. By including triangle as random intercept, we allow the trials to have a different effect on pitch. Since we are not looking at a development, we did not include any random slopes. The participants are paired across diagnosis, therefore gender and age is not included as fixed effects as this is already accounted for in the data. 

Comparison of their performance: 
In order to assess the performance of the model we used the model to predict diagnosis. With the predict() function the outcome is in log odds. We used the inv.logit() to obtain pobabilities for intepretation reasons. 
We used the predictions to create a ROC curve and examined the area under the curve and the actual plot. The area under the curve is 0.62 with 95% confidence intervals: 0.59 - 0.65.

Confusion Matrix:

Prediction Control Schizo
   Control     300    119
   Schizo      364    556
                                         
               Accuracy : 0.6393         
                 95% CI : (0.6129, 0.665)

            Sensitivity : 0.8237         
            Specificity : 0.4518         
         Pos Pred Value : 0.6043         
         Neg Pred Value : 0.7160
         
         
The accuracy of the model is 0.64, which is not sufficient enough to actually rely on in the real world. Investigating the model further looking at sensitivity and specificity we can see that the model has a quite high sensitivity (0.82), but a low specificity (0.45). The sensitivity of the model is the rate that the event of interest is predicted correctly for all samples having the event. Because of the high sensitivity in this model, there is a good chance of diagnosing the subject as schizophrenic, if the subject is in fact schizophrenic. But specificity is rather low, this indicates that this model would often diagnose subjects as scizophrenics, when they were in fact controls.
(The specificity is defined as the rate that nonevent samples are predicted as nonevents).


```{r}
plot(rocCurve, legacy.axes=TRUE)
auc(rocCurve)
ci(rocCurve)

```

We cross-validated the model to get a more realistic idea of the performance of the model and got the following confusion matrix:

Prediction Control Schizo
   Control     246    202
   Schizo      418    473
                                          
               Accuracy : 0.537           
                 95% CI : (0.5098, 0.5639)

            Sensitivity : 0.7007          
            Specificity : 0.3705          
         Pos Pred Value : 0.5309          
         Neg Pred Value : 0.5491          

From this we can see that the model actually performs worse than originally thought with an accucary of 0.54 and worse specificity (0.37) and sensitivity (0.70).

#QUESTION 2: Feature selection

We cross-validated a models with each acoustic feature:

glmer(Diagnosis ~ acoustic feature + (1|Subject) + (1|Triangle), family = "binomial")

The outcome of this was as follows:

[1] "z.inter quartile range .95 has an accuracy of: 0.565 a sensitivity of: 0.776 and a specificity of: 0.351"
[1] "z.Mean has an accuracy of: 0.499 a sensitivity of: 0.399 and a specificity of: 0.601"
[1] "z.sd has an accuracy of: 0.557 a sensitivity of: 0.778 and a specificity of: 0.333"
[1] "z.q05 has an accuracy of: 0.538 a sensitivity of: 0.425 and a specificity of: 0.652"
[1] "z.q25 has an accuracy of: 0.523 a sensitivity of: 0.415 and a specificity of: 0.633"
[1] "z.q50 has an accuracy of: 0.483 a sensitivity of: 0.39 and a specificity of: 0.578"
[1] "z.q75 has an accuracy of: 0.476 a sensitivity of: 0.447 and a specificity of: 0.505"
[1] "z.q95 has an accuracy of: 0.453 a sensitivity of: 0.41 and a specificity of: 0.497"
[1] "z.inter quartile range .75 has an accuracy of: 0.532 a sensitivity of: 0.686 and a specificity of: 0.375"
[1] "z.MAD has an accuracy of: 0.559 a sensitivity of: 0.721 and a specificity of: 0.395"
[1] "z.CoefVar has an accuracy of: 0.591 a sensitivity of: 0.799 and a specificity of: 0.381"

From this we will conclude that inter quartile range 0.95 (accuracy: 0.57, sensitivity: 0.776, specificity: 0.351), sd (accuracy: 0.55, sensitivity: 0.778, specificity: 0.333), MAD (accuracy: 0.56, sensitivity: 0.721, specificity of: 0.395), Coefficient of variance (accuracy: 0.59, sensitivity: 0.799, specificity: 0.381) are the best predictors of diagnosis. The models accuracy score are similar and by eyeballing the ratio between sensitivity and specificity, we see that these are also similar across the models. From this it is difficult to select one model as being better than the others.  


QUESTION 3



###RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

In conclusion, Diagnosis cannot be predicted from acoustic features from voice pitch nor pause. Since none of the model perform better than an accuracy of ~ .60.
However, the sensitivity of the model however is rather high ~.80. This assures that there is a good chance that if a subject is scizophrenic, this subject will actually get the diagnosis; shizophrenic.


The best model:
glmer(duration of pause * number of pause + coe.var * MAD * iqr2575 + (1 | Subject) + (1 | Triangle))







### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?
