---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Laurits Dixen"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(caret)
library(Metrics)
library(lmerTest)
library(boot)
library(pROC)

#loading the data
data = read.csv("schizoAcousticData.csv")
# Making the subjects paired again
data$Subject = ifelse(data$Diagnosis==0, data$Subject-1000, data$Subject) 

#making list for feature names for later use
features = colnames(data)[6:16]

```

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

### Question 1: Can you diagnose schizophrenia from pitch range only? If so, how well?

Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

```{r}

#turning pitch range into z-score (needed for the model to converge) (can also be done like this: data$z.qr95 = scale(data$qr95, center = TRUE, scale = TRUE))
data$z.qr95 = (data$qr95 - mean(na.omit(data$qr95))) / sd(na.omit(data$qr95))

#create a model with diagnosis predicted by pitch range (0.5 to 0.95 quantiles) and random effects
model = glmer(Diagnosis ~ z.qr95 + (1|Triangle) + (1|Subject), family = "binomial",data = data)
summary(model)

#creating extra columns with all features z-scored 
for (f in features){ #loop that runs through the list of features and paste it into the calculation
  eval(parse(text = paste("data$z.",f," = (data$",f," - mean(na.omit(data$",f,"))) / sd(na.omit(data$",f,"))", sep="")))
}

#making list with z-scored feature names for later use
zfeatures = c(colnames(data)[32:41],"z.qr95")

#add a column with predictions in the unit "probabilities" (with inv.logit)
data$predictions = inv.logit(predict(model, newdata = data, allow.new.levels=TRUE))

#set threshold for diagnosing at 0.5
data$predictions[data$predictions > 0.5] = "Schizo"
data$predictions[data$predictions <= 0.5] = "Control"

#align labels of diagnosis
data$Diagnosis[data$Diagnosis == 1] = "Schizo"
data$Diagnosis[data$Diagnosis == 0] = "Control"

#create confusion matrix to get performance measure
confusionMatrix(data = as.factor(data$predictions), reference = as.factor(data$Diagnosis), positive = "Schizo")

#load predictions again to get numeric data
data$predictions = inv.logit(predict(model, newdata = data, allow.new.levels=TRUE))

#ROC curve
rocCurve <- roc(response = data$Diagnosis, predictor= data$predictions)
#get area under the curve
auc(rocCurve)
#95% confidence intervals
ci(rocCurve)
#plot the curve
plot(rocCurve, legacy.axes=TRUE)

```


Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

```{r}
#make diagnosis numeric again
data$Diagnosis[data$Diagnosis == "Schizo"] = 1
data$Diagnosis[data$Diagnosis == "Control"] = 0
data$Diagnosis = as.numeric(data$Diagnosis)

#cross validation with 5 folds over the pitch range model
folds = createFolds(y = unique(data$Subject), k=5, list = TRUE, returnTrain = FALSE)
folds = lapply(folds, function(x) unique(data$Subject)[x])
testSet = NULL
for (f in folds) {
  test = filter(data, Subject %in% f)
  train = filter(data, !Subject %in% f)
  
  tempModel = glmer(Diagnosis ~ z.qr95 + (1|Subject) + (1|Triangle), family = "binomial",data = train)
  
  test$predictions = inv.logit(predict(tempModel, newdata = test, allow.new.levels=TRUE))

  # Creating the full dataframe of predictions (can also be done like this: "predictions[data$Subject %in% f] = test$predictions", need to create emtpy vector before loop, like this "predictions = rep(NA, nrow(data))")
  if (!is.null(testSet)){
    testSet = rbind(testSet,test)
  } else {
    testSet = test
  }  
}

#set threshold for diagnosing at 0.5 and change label
testSet$predictions[testSet$predictions > 0.5] = "Schizo"
testSet$predictions[testSet$predictions <= 0.5] = "Control"

#align labels of diagnosis
testSet$Diagnosis[testSet$Diagnosis == 1] = "Schizo"
testSet$Diagnosis[testSet$Diagnosis == 0] = "Control"

#create confusion matrix to get performance measures
confusionMatrix(data = as.factor(testSet$predictions), reference = as.factor(testSet$Diagnosis), positive = "Schizo")

```


N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

### Question 2 - Which single acoustic predictor is the best predictor of diagnosis?

```{r}
#make diagnosis numeric again
data$Diagnosis[data$Diagnosis == "Schizo"] = 1
data$Diagnosis[data$Diagnosis == "Control"] = 0
data$Diagnosis = as.numeric(data$Diagnosis)

#cross validate the models (one for each acoustic predictor)
folds = createFolds(y = unique(data$Subject), k=5, list = TRUE, returnTrain = FALSE)
folds = lapply(folds, function(x) unique(data$Subject)[x])
testSet = NULL 
for (f in folds) {
  test = filter(data, Subject %in% f)
  train = filter(data, !Subject %in% f)
  
  for (zf in zfeatures) { #loop that runs through the list of z.features  
    eval(parse(text=paste('tempModel = glmer(Diagnosis ~ ',zf,' + (1|Subject) + (1|Triangle), family = "binomial",data = train)', sep='' )))
    eval(parse(text=paste('test$pred',zf,' = inv.logit(predict(tempModel, newdata = test, allow.new.levels=TRUE))',sep='' )))
    eval(parse(text=paste('test$pred',zf,'[test$pred',zf,' > 0.5] = "Schizo"',  sep=''))) #set threshold
    eval(parse(text=paste('test$pred',zf,'[test$pred',zf,' <= 0.5] = "Control"',sep='')))
  }
  
  # Creating the full dataframe of predictions
  if (!is.null(testSet)){
    testSet = rbind(testSet,test)
  } else {
    testSet = test
  }  
}

#change labels for diagnosis
testSet$Diagnosis[testSet$Diagnosis == 1] = "Schizo"
testSet$Diagnosis[testSet$Diagnosis == 0] = "Control"

#make confusion matrix to get performance measures
for (zf in zfeatures){ #loop that runs through all the z.features 
  eval(parse(text=paste('matrix = confusionMatrix(data = as.factor(testSet$pred',zf,'), reference = as.factor(testSet$Diagnosis), positive = "Schizo")',sep = '')))
  print(paste(zf,"has an accuracy of:",round(matrix$overall[1],3))) #print the accuracy score for all the models
}



```


### Question 3 - Which combination of acoustic predictors is best for diagnosing schizophrenia?

Now it's time to go wild! Use all (voice-related) variables and interactions you can think of. Compare models and select the best performing model you can find.

Remember:
- Out-of-sample error crucial to build the best model!
- After choosing the model, send Malte and Riccardo the code of your model
```{r}
read_pitch <- function(filename) {
    # read data
  data <- read.delim(file = paste(path, filename, sep = ""), sep = "\t", header = TRUE) 
  soundname = filename
    # parse filename
  meta = str_extract_all(filename, "\\d+")
  study = as.numeric(meta[[1]][1])
  Diagnosis = as.numeric(meta[[1]][2])
  Subject = as.numeric(meta [[1]][3])
  triangles = as.numeric(meta[[1]][4])
  
    # extract descriptors
  mean_f0 = round(mean(data$f0), digits = 2)
  sd = sd(data$f0)
  quantiles = quantile(data$f0, probs = c(0.05, 0.25, 0.50, 0.75, 0.95))
  q05 = quantiles[[1]]
  q25 = quantiles[[2]]
  median = quantiles[[3]]
  q75 = quantiles[[4]]
  q95 = quantiles[[5]]
  iqr2575 = q75-q25
  coe.var = coefficient.variation(sd(data$f0), mean(data$f0))
  MAD = mad(data$f0) #average distance from something...
  
    # combine all this data
  output_list = data.frame(Subject, Diagnosis, study, triangles, mean_f0, sd, q05, q25, median, q75, q95, iqr2575, coe.var, MAD,soundname)
  
  return(output_list)
}
path = "/Users/lauritsdixen/R-stuff/Assignment 2/cogsci-methods-3-assignments-master/Assignment3_VoiceSchizo/Pitch/"

pitch_data = list.files(path = path) %>%
    purrr::map_df(read_pitch)

extra = read.delim("Articulation.txt", sep=",")

fullData = full_join(pitch_data,extra, by = "soundname")
class(pitch_data$soundname)
class(extra$soundname)
```


```{r}

#Some libraries for special functions
library(rje)
library(combinat)
# Defining possible variables
variables <- c('z.qr75','z.sd','z.MAD','z.CoefVar')




models = data.frame(matrix(ncol=4),nrow=0)
colnames(models) = c("variables", "accuracy", "sensi", "speci")

folds = createFolds(y = unique(data$Subject), k=5, list = TRUE, returnTrain = FALSE)
folds = lapply(folds, function(x) unique(data$Subject)[x])
# Creating a powerset (set of all possible subsets) for the permutations
PowerSet <- powerSet(variables)
for (set in PowerSet[2:16]) {
  # Reformating the set
  currentSet <- c()
  for (e in set) {
    currentSet <- c(e,currentSet) }
  # Putting the variables in a string seperated by a '+'
  collapsed <- paste(currentSet, collapse ="+")
  print(collapsed)
  # Creating the command in string format
  testParsing <- paste('glmer(Diagnosis ~ ',collapsed,' + (1|Subject) + (1|Triangle), data = train, family = "binomial")',sep = '')
  predictors = rep(NA, nrow(data))
  # CV folds
  for (f in folds) {
    test = filter(data, Subject %in% f)
    train = filter(data, !Subject %in% f)
    # Training the model
    model <- eval(parse(text=testParsing))
    # Creating predictions
    predictors[data$Subject %in% f] = as.numeric(inv.logit(predict(model, newdata = test, allow.new.levels=TRUE)))
  }
  predictors = as.numeric(predictors)
  predictors = ifelse(predictors > 0.5, 1, 0)
  matrix = confusionMatrix(data = as.factor(predictors), reference = as.factor(data$Diagnosis),positive = "1")
  models = rbind(models, c(collapsed,matrix$overall[1],matrix$byClass[1],matrix$byClass[2]))
  }
}
```


### Question 4: Properly report the results

METHODS SECTION: how did you analyse the data? That is, how did you extract the data, designed the models and compared their performance?

RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?

### Bonus question 6

Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them.
